---
title: "STAT 474 &ndash; Techniques for Large Data Sets"
subtitle: "Fall 2018"
date: "September 26, 2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, message=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(eval=FALSE)
library(mongolite)
example <- mongo(collection = "example", 
                 url = "mongodb://localhost")
example$remove('{}')
```

# Introduction to MongoDB - Review

- What is MongoDB? How can we discribe the MongoDB?

--

- We have been working with several MongoDB functions: `insert()`, `find()`, `count()`, `distinct()` and `iterate()`. Can you discribe the functionality of each function?

--

- MongoDB query syntax - following JSON structure: 

  - `<field>:<value> `
  
  - `<field>: { <operator>: <value> }`

  - Write a query to find all the robberies or theft incidents in 2018 from Chicago's crime data, assuming that the handler for the collection is `crimes`.
  
---
# Basic CRUD Operations in MongoDB

- **CRUD** stands for *creating*, *reading*, *updating* and *deleting*.

- We have worked with *creating* via `insert()` function.

  - Accepts data frame or JSON strings (one string for each document)
  
  - Atomicity: single document insertion is atomic, while multiple-document insertion is not.
  
```{r}
example <- mongo(collection = "<YOUR OWN COLLECTION NAME>", 
        url = "mongodb://<USERNAME>:<PASSWORD>@<IP ADDRESS>/test")

queries <- c('{"name" : "Peter", "age" : 99, "loves": ["apple"]}',
      '{"name" : "Ruth", "age" : 93, "loves": ["apple", "sugar"]}',
      '{"name" : "Jack", "loves" : ["bananas"]}');
example$insert(queries)
example$find()
```

---
# Basic CRUD Operations in MongoDB

- We have worked with *reading* via `find()` function.

  - `fields=` option allows selection of listed fields
  
  - `sort=` option alows sorting
  
  - *Activity:* What happen when you change the number 1? Try other numbers, including -1. Comment and take note to yourself about the sorting feature.
  
```{r}
example$find('{}', sort = '{"age" : 1}')
```

  
---
# Basic CRUD Operations in MongoDB

- *Deleting* operation is straightforward

  - To delete the whole collection, use `drop()` as in `crimes$drop()`. In practice, there is very few time when we want to delete a collection.
  
  - More often, we need to remove one/many documents: use `remove(query = '{}')`.
  
---
# Updating in MongoDB

- The function `update()` takes two arguments: the selector (i.e. WHERE in SQL) to use and what updates to apply to fields.

```{r}
example$update('{"name" : "Peter"}', '{"$set" : {"age" : 90}}')
```

- *Activity:* Verify the effect of updating command. What happens if you forget the `$set` operator?

---
# Adding value to an array

```{r}
example$update('{"name" : "Peter"}', 
               '{"$push" : {"loves" :  "candy"}}')
```

- *Activity:* What happens if we mistakenly use `$set` instead? 
---
# Upserts

- An upsert updates the document if found or inserts it if not.

```{r}
example$update('{"name": "Bob"}','{"$inc":{"age" : 1}}', upsert = T)
```

- *Activity:* What do you observe when running the above query? Comment on the effect of `$inc`?

---
# Updating in MongoDB

- I'd like to add a field `vacinated` (boolean type) with the value of `TRUE` for all documents. What do you observe when running the below query?

```{r}
example$update('{}', '{"$set" : {"vacinated" : true}}')
```

--

- Use the option `multiple = TRUE` for multiple updates. 

- *Activity:* Try the `multiple = TRUE` option with above query and verify the expected result.

