---
title: "STAT 474 &ndash; Techniques for Large Data Sets"
subtitle: "Fall 2018"
date: "November 5, 2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(eval=FALSE)
library(stringr)
library(sentimentr)
```

# Text mining - Review

- Collection of methods and techniques to extract, summarize, and analyze useful information from text data

--

- Text mining project:
  
  - Text preprocessing: cleaning up, tokenization, normalization, etc.
  
  - Feature generation, e.g., word frequencies
  
  - Data mining. e.g., classification, regression

--

- **What else have we done?**

---
# What is Sentiment?

- Sentiment = feelings

  - Also means attitudes, emotions, opinions
  
  - Sentiments are subjective impressions, not facts.
  
--

- When working with sentiments, a binary opposition in opinions is assumed

  - For/against, like/dislike, good/bad, etc.
  
  - A few projects work with other types of emotions, but not common.

---
# What is Sentiment Analysis?

- Using NLP (natural language process), statistics, or machine learning methods to extract, identify, or otherwise characterize the sentiment content of a text unit

- Sometimes referred to as **opinion mining**, although the emphasis in this case is on extraction

--

- Questions sentiment analysis might ask:

  - Is this product review positive or negative?
  
  - Is this customer email satisfied or dissatisfied?
  
  - How have bloggers' attitudes about the president changed since the election?

---
# Other related taskes for sentiment analysis
  
- Information extraction (discarding subjective information)

- Question answering (recognizing opinion-oriented questions)

- Summarization (accounting for multiple viewpoints)

- "Flame"/Insulting detection

- Identifying child-suitability of videos based on comments

- Bias identification in news sources

---
# An application in Business Analytics

- Question: "Why aren't consumers buying our laptop?"

- We know the concrete data: price, specs, competition, etc.

--

- We want to know subjective data: "the design is tacky", "customer service was horrible"

- Misperceptions are also important, e.g., "updated drivers aren't available" (actually, they are!!)

--

- It is very difficult to survey customers who **didn't** by the company's laptop

- Instead, you could use sentiment analysis to

  - search the web for opinions and reviews of this and competing laptops: blogs, amazon, tweets, etc.
  
  - create condensed versions or a digest of consensus points.
---
# Challenges in sentiment analysis

- People express opinions in complex ways.

- In opinion texts, lexical content alone can be misleading

- Intra-textual and sub-sentential reversals, negation, topic change common

- Rhetorical devices/modes such as sarcasm, irony, implication, etc. 

---
# Example - a fictitious letter to a hardward store

> “Dear &lt;hardware store&gt;

> Yesterday I had occasion to visit &lt;your competitor&gt;. They had an excellent selection, friendly and helpful salespeople, and the lowest prices in town.

> You guys suck.

> Sincerely,”  

--

- Humans are subjective creatures and opinions are important. Being able to interact with people on that level has many advantages.

---
# Where to start

- Decide on the building blocks of sentiment expression. 

  - Many possibilities: users, texts, sentences, words, tweets/updates
  
--

- It is common to use words as building blocks

  - Short phrases may be just as important (or moreso) as words, e.g., "lowest prices", "high quality"
  
--

- There seems to be **some** relation between positive words and positive sentiments

  - Can we come up with a set of keywords to identify sentiments?
  
---
# Keyword methods

- By hand: labor intensive. Amazon's Mechanical Turk for NLP annotation

  - Roughly $1 for 1,000 labels
  
  - 5 non-expert annotators achieve equivalent accuracy to 1 expert annotator
--

- Data-driven methods can be used to generate keyword lists

  - model better than human-generated lists (might due to the volume)
  
  - Comments, tweets with emoticons/smileys
  
  - Reviews (Amazon, Yelp) with product/item ratings.

--

- Sentiment-oriented data set is highly domain sensitive and very difficult to create/collect.
  
---
# Availability of sentiment-oriented data sets

- Many linguists open their databases for public use.

  - `tidytext` package has 4 different lexicons for sentiment analysis 
  
  - **What are their differences?**
  
--

The three general-purpose lexicons:

- `AFINN` from [Finn Arup Nielsen](http://www2.imm.dtu.dk/pubdb/views/publication_details.php?id=6010): with scores from -5 to 5.

- `bing` from [Bing Liu and collaborators](https://www.cs.uic.edu/~liub/FBS/sentiment-analysis.html): with positive/negative annotation.

- `nrc` from [Saif Mohammad and Peter Turney](http://saifmohammad.com/WebPages/NRC-Emotion-Lexicon.htm): many different types of emotions

---
# Demo on Sentiment Analysis

- Examine how the sentiment changes across the novels.

- The narrative arc is positive if the total "score" is positive, and is negative if the total "score" is negative

  - Difference in counts of positive and negative words (using `bing`)
  
  - Total sum of sentiment score when using `affin`
  
- Define a book section as 80 lines of text

  - Too small sections might not have enough words (not all words have sentiment scores)
  
  - Too long sections might wash out narrative structure.
---
# Issues with keyword methods

- Words may not be enough

  - *ridiculous comedy*  versus *ridiculus drama*
  
  - *cheap construction* versus *cheap deal*
  
--

- We might want to assign sentiment scores to certain kinds of phrases, e.g "high quality", etc.

- Binary sentiment don't capture nuance

--

- Negations (in general **valence shifters**) changes the sentiment score of a word

  - Example: "I am not very happy"
  
---
# Looking at units beyond just words
  
- Some sentiment analysis algorithms look beyond only unigrams (i.e. single words) to try to understand the sentiment of a sentence as a whole.

- Token can be sentences or even chapters.

  - It is harder to clean up the text.
  
--

- R packages: `coreNLP`, `cleanNLP`, `sentimentr`, etc.

- Example of `sentimentr`

```{r, eval=TRUE}
sentiment('I am not very happy')
```

- It takes more computational power (and time) to analyse with bigger token structure.