---
title: "STAT 474 &ndash; Techniques for Large Data Sets"
subtitle: "Fall 2018"
date: "September 12, 2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(eval=FALSE)
```

# What we covered on Monday?

--

- Set operations in SQL: `INTERSECT`, `UNION`, and `UNION ALL`

--

- A subquery is a `SELECT` statement that is embedded in a clause of another `SELECT`.
  
--

  - Find headquarter cities of companies having factories in both US *and* China.

```{r}
Company (name, hq_city)
Product (pname, maker, factory_loc)
```

```{r}
SELECT DISTINCT hq_city
FROM   Company, Product
WHERE  maker = name 
       AND name IN ( #<<
		SELECT maker
	  	FROM   Product
	  	WHERE  factory_loc = ‘US’)
	 AND name IN ( #<<
		SELECT maker
	  	FROM   Product
	  	WHERE  factory_loc = ‘China’)
```

---
# Subqueries

- Can build powerful statements out of simple ones by using subqueries.

- You can place the subsquey in a number of SQL clauses: `WHERE`, `HAVING`, `FROM`.

- In the syntax, comparison operator falls into two classes:

  - Single-row operators ( >, =, >=, <, <=, <>)

  - Multiple-row operators (with `IN`, `ANY`, `ALL`)

--

Example: `Product(name, price, category, maker)`

```{r}
SELECT name
FROM   Product
WHERE  price > ALL( #<<
	SELECT price
     FROM   Product
     WHERE  maker = ‘Gizmo-Works’);
```

---
# What is wrong with this statement?

```{r}
Product(name, price, category, maker)
```

```{r}
SELECT name
FROM   Product
WHERE  price = ( 
	SELECT MIN(price)
     FROM   Product);
```

---
# Nested queries as alternatives to `INTERSECT`

- *Note:* `INTERSECT` might **not** be available in some DBMSs!

```{r}
(SELECT R.A, R.B
 FROM   R)
INTERSECT
(SELECT S.A, S.B
 FROM   S)
```

--

```{r}
SELECT R.A, R.B
FROM   R
WHERE EXISTS(
   	SELECT *
   	  FROM S
   	WHERE R.A=S.A AND R.B=S.B)
```

---
# Correlated Queries
### Using External Vars in Internal Subquery

> Find movies whose title appears more than once.

```{r}
Movie(title, year, director, length)
```

--

```{r}
SELECT DISTINCT title
FROM   Movie AS m #<<
WHERE  year <> ANY( 
		SELECT  year
     	FROM    Movie  #<<
     	WHERE  title =  m.title) #<<
```

- *Note:* the scoping of the variables!

---
# Complex Correlated Query

> Find products (and their manufacturers) that are more expensive than all products made by the same manufacturer before 1972

```{r}
Product (name, price, category, maker, year)
```

<br>

.center[ **Let's pause and attempt to write such query**]

--

<br>

```{r}
SELECT DISTINCT  x.name, x.maker
FROM   Product AS x
WHERE  x.price > ALL(
		SELECT y.price
          FROM   Product AS y
          WHERE  x.maker = y.maker 
		   AND y.year < 1972)
```

**Note:** Complex correlated query can be very powerful (also much harder to optimize)

---
# Nested Query vs. `GROUP BY`

```{r}
Author (login, name)
Wrote (login, url)
```

> Find authors who wrote 10 or more documents.

--

*Attempt 1: with nested queries*

```{r}
SELECT DISTINCT Author.name
FROM   Author
WHERE  COUNT(
	SELECT Wrote.url
	FROM   Wrote
	WHERE  Author.login = Wrote.login) > 10
```

---
# Nested Query vs. `GROUP BY`

```{r}
Author (login, name)
Wrote (login, url)
```

> Find authors who wrote 10 or more documents.

--

*Attempt 2: without nested queries*

--

```{r}
SELECT   Author.name
FROM     Author, Wrote
WHERE    Author.login = Wrote.login
GROUP BY Author.name
HAVING   COUNT(Wrote.url) > 10
```

**Question:** Do we need `DISTINCT` in the above query?

---
# Nested Query vs. `GROUP BY`

Which way is more efficient?

--

- Attempt 1 - with nested queries: How many times do we do a SFW query over all of the Wrote relations?

--

- Attempt 2 - with `GROUP BY`: How about when written this way?

--

.center[With `GROUP BY`, queries can be much more efficient!]

---
# Views

- Subqueries are powerful, but make the queries easy to become complex and hard to "debug".

- It is always better to break down into intermediate tables.

- *Drawback:* The new table **uses up storage**, even though it is recomputable from other tables.

--

- Stored output tables are some time problematic as it might be **out of date.**

--

- SQL provides a notion of logical **views**: basically named queries that are re-evaluated upon each reference.

- The syntax is straightforward:

```{r}
CREATE VIEW <name> AS
<SELECT statement>;
```

---
# Using views in SQL

- Example: [http://sqlfiddle.com/#!9/0fb0bd](http://sqlfiddle.com/#!9/0fb0bd)

--

- Queries can use **view** as if it were a base table

--

- View always up-to-date

  - (Re-)evaluated whenever a query uses the view
  
  - Keeping it up-to-date is responsibility of the DBMS and not the users

--

- `DROP VIEW` command: dispose of a view

---
# Updating a View

- What if an update is applied to a view as if it were a base table? (see [http://sqlfiddle.com/#!9/0fb0bd](http://sqlfiddle.com/#!9/0fb0bd))

--

- Views are updateable. You can use the `INSERT` or `UPDATE` to add or modify rows in the base table. `DELETE` applied on view can remove rows in the base table

--

- Not all views are updatable.

  - What if view contain aggregate functions?
  
  - What if the view is defined as a `UNION` of two tables?
  
---
# Too many views

Views can help:

- Simplify queries

- Make complex queries more readable

- Share "sql programs" with others

--

<br>

Problems:

- Creating a new view for each (exploratory) query will result in a lot if views! Example: `temp1`, `temp1_joey`, `temp1_joey_fixed`,...

- Often you will not be allowed to create views 

---
# Common Table Expressions (`WITH`)

- Think of these as a view that exists only during query

- Syntax:

```{r}
WITH <name> [(renamed columns)] AS
  (<SELECT statement>)
[ , <name2> AS (<SELECT statement>)...]
```

- Can define one or more CTEs in `WITH` clause.

- Level of support for CTEs varies across different DBMSs. Microsoft SQL Server has been supported CTEs for a long time, while MySQL only starts supporting with its new version 8.0.