---
title: "Simple Web Sccrapping with R"
output: html_notebook
---

In R, the package `rvest` makes it easy to scrape (or harvest) data from html web pages. The package was inspired by libraries like [beautiful soup](https://www.crummy.com/software/BeautifulSoup/), which was written in Python. It is designed to work with [magrittr](https://github.com/tidyverse/magrittr) so tgat you can express complex operations as elegant pipelines composed of simple, easily understood pieces. Install it with


```{r}
install.packages("rvest")
library(rvest)
```

# Examining Solar Energy Usage:

Suppose I am interested in looking at photovoltaic output for California. This data is aggregated on the web here: https://www.energy.ca.gov/almanac/renewables_data/solar/index.php

To read the html page from the Internet:

```{r}
page <- read_html("https://www.energy.ca.gov/almanac/renewables_data/solar/index.php")
```

Then, we extract all available tables from the web page

```{r}
tables <- html_table(page)
```

This returns a list of tables.

```{r}
length(tables)
```

We can look at each table individually,

```{r}
head(tables[[1]])
```

or all of them at once

```{r}
tables
```

If we return to the website here:

[http://www.energy.ca.gov/almanac/renewables_data/solar/index.php](http://www.energy.ca.gov/almanac/renewables_data/solar/index.php)

we see that this reflects the content on the website.

# Cleaning the data

When extracting data from the web with `rvest` we are likely to need to do substantial cleaning. For example, let's look at the first table

```{r}
head(tables[[1]])
str(tables[[1]])
```

First, we save the body of table into a different object and fix the header

```{r}
pv <- tables[[1]][-1,]
names(pv) <- tables[[1]][1,]
head(pv)
```

How's about the end of the table

```{r}
tail(pv)
```

Since a tidy data should have row represent an observational unit, the last row should be removed

```{r}
pv <- pv[-nrow(pv), ]
```

While it seems that we have finished cleaning up the data, it is important to check the data types of columns:

```{r}
str(pv)
```

We observe that numeric columns are all coded as `chr`, indicating a *string* type. 

```{r}
pv$`Capacity (MW)` <- as.numeric(pv$`Capacity (MW)`)
pv$`Net MWh` <- as.numeric(pv$`Net MWh`)
```

Oopses! There is an error message when working with `Net Mwh`. Let's re-run everything, but the last step. You will see that column `Net Mwh` has numbers with comma separatord at the thousands and `as.numeric()` function can not handle that. To work around the issue, we first remove comma and then use `as.numeric()` function again

```{r}
pv$`Net MWh` <- as.numeric(gsub(",", "", pv$`Net MWh`))
```

Final check:

```{r}
head(pv)
```

Once we've done with cleaning the data set, we should save it to disk for further analysis

```{r}
write_csv(pv, "pv.csv")
```

**Tasks:** Obtain and clean up the Solar Photovoltaic data set from the website

*Note:* 
- Some of the company names and EIA Plant IDs are missing. They should be coded as `NA` in R. 
- When dealing with text, the function `tolower()` and `toupper()` are very useful. They convert characters into their lower/upper cases.