---
title: "Working with JSON in R"
output: html_notebook
---

A few R packages such as `jsonlite`, `rjson`, `RJSONIO`, provide basic JSON functionality for R users and other packages (e.g., `mongolite`). We will use the package `jsonlite` for the demo (it is also part of the `tidyverse`)

```{r}
# The package is needed to be installed once. Uncomment if this is your first use.
# install.packages("jsonlite")
library(tidyverse)
library(jsonlite)
```


The first example is based on the simple JSON data

```{r}
data <- '[
{
    "Prof": "Gonzalez",
    "Classes": [
      "CS186", 
      { "Name": "Data100", "Year": [2017,2018] }
    ],
    "Tenured": false
  },
  {
    "Prof": "Nolan",
    "Classes": [
      "Stat133", "Stat153", "Stat198", "Data100"
    ],
    "Tenured": true
  }
]'
data
```

*Note:* The `\n` in the string `data` is a newline character, which is used to signify the end of a line of the text and the start of the new one.

To display a string without its control characters (e.g., newline characters), we can use the function `cat()` (similar to `cat` command we encountered when working with Hadoop filesystem)

```{r}
cat(data)
```


To read an JSON object, we use `fromJSON`:

```{r}
json_data <- fromJSON(data)
json_data
```

**Question:** Compare the display of `json_data` to the original structure and comment on the differences.

Examine the structure of the parsed JSON object:

```{r}
str(json_data)
```

The overall structure of `json_data` is a data frame, which allows use to access individual columns via indexing or `$`

```{r}
json_data$Prof
json_data[ , -2]
```

Accessing data in `list` format is tricker and requires using the list indexing `[[ ... ]]`. For example, to read data about the `Data 100` class from Prof Gonzales

```{r}
json_data$Classes[[1]][[2]]
json_data$Classes[[1]][[2]][["Year"]]
```

**Question:** Can we list the courses Prof. Nolan taught?

