---
title: "Lab 2 - Multi-table Queries"
output: html_notebook
---


We start with loading those packages into the current session:
```{r, message=FALSE}
library(DBI)
library(RSQLite)
```

Next, we will treat the memory as our database:

```{r}
con <- dbConnect(SQLite(), dbname=":memory:")
```

Run the following codes. They are setting up the tables that we will use in this lab.

```{r, message=FALSE}
dbSendQuery(con, 'DROP TABLE IF EXISTS R;')
dbSendQuery(con, 'DROP TABLE IF EXISTS S;')
dbSendQuery(con, 'DROP TABLE IF EXISTS T;')
dbSendQuery(con, 'CREATE TABLE R (A int);')
dbSendQuery(con, 'CREATE TABLE S (A int);')
dbSendQuery(con, 'CREATE TABLE T (A int);')
for (i in 1:6) {
  query <- paste0("INSERT INTO R VALUES (", i, ");")
  dbSendQuery(con, query)
}
for (i in seq(1, 10, 2)) {
  query <- paste0("INSERT INTO S VALUES (", i, ");")
  dbSendQuery(con, query)
}
for (i in seq(1, 11, 3)) {
  query <- paste0("INSERT INTO T VALUES (", i, ");")
  dbSendQuery(con, query)
}
```


```{r, message=FALSE}
dbSendQuery(con, 'DROP TABLE IF EXISTS product;')
dbSendQuery(con, 'DROP TABLE IF EXISTS company;')
dbSendQuery(con,'pragma foreign_keys = ON; -- WARNING by default off in sqlite)')
dbSendQuery(con, '
CREATE TABLE company (
    cname VARCHAR PRIMARY KEY, -- company name uniquely identifies the company.
    stockprice MONEY, -- stock price is in money 
    country VARCHAR); -- country is just a string
')
dbSendQuery(con,'
INSERT INTO company 
VALUES (\'ToyWorks\', 25.0, \'USA\'),
       (\'ToyFriends\', 65.0, \'China\'),
       (\'ToyCo\', 15.0, \'China\');
')

dbSendQuery(con,'
CREATE TABLE product(
       pname VARCHAR, -- name of the product
       price MONEY, -- price of the product
       category VARCHAR, -- category
       manufacturer VARCHAR, -- manufacturer
       PRIMARY KEY (pname, manufacturer),
       FOREIGN KEY (manufacturer) REFERENCES company(cname));
')
dbSendQuery(con,'
INSERT INTO product 
VALUES (\'Pikachu\', 19.99, \'Toy\', \'ToyWorks\'),
       (\'Pikachu\', 19.99, \'Toy\', \'ToyFriends\'),
       (\'Pokeball\', 29.99, \'Electronic\', \'ToyCo\'),
       (\'Bulbasaur\', 149.99, \'Toy\', \'ToyFriends\'),
       (\'Charizard\', 203.99, \'Toy\', \'ToyCo\'),
       (\'PokeCamera\', 19.99, \'Electronic\', \'ToyWorks\');
')
```

**Question 1:**

For three tables $R,S,T$ that only have one attribute $A$:

- R = {1,2,3,4,5}
- S = {1,3,5,7,9}
- T = {1,4,7,10}

Can you write a query to select $R \cap (S \cup T)$- in other words elements that are in $R$ and either $S$ or $T$?

Write your query here:
```{r}
dbGetQuery(con,'

')
```

Now test your query above for the case where $S = \emptyset$- what happens and why?

Execute the below, then re-run your query above

```{r}
dbSendQuery(con, 'DELETE FROM S;)')
```



**Question 2:**

- Schema is the same as before

> Product (pname, price, category, manufacturer)
> Company (cname, stockPrice, country)

Write your query to find all categories of products that are made by Chinese companies

```{r}
dbGetQuery(con,'

')
```