---
title: "Lab 1 - Basic SQL"
output: html_notebook
---

# Part 1: Schemas & table creation

Before we actually get into basic SQL queries (asking questions of data in tables), we'll look at some of the basics about how to create tables.

*NOTE:* Make sure to have a copy of the database file, `dataset_1.db`, from the last lecture downloaded and in this directory for the below to work!

We start with loading those packages into the current session:
```{r, message=FALSE}
library(DBI)
library(RSQLite)
```

Next, we will load an SQLite database stored as a file as follows:

```{r}
con <- dbConnect(SQLite(), dbname="dataset_1.db")
```


Recall that the database we just loaded has one table, `precipitation_full`, having the following schema:

- state_code
- station_id
- year
- month
- day
- hour
- precipitation
- flag_1
- flag_2

Each tuple in this table describes one hour of rainfall (`precipitation-` in hundredths of an inch) at one station (`station_id`) in one state (`state_code`). Note that tuples with `hour=25` record the total rainfall for that day, and that we can ignore the values of attributes `flag_1` and `flag_2` for now.

Now, however, let's see how to view the schema of existing tables on your own; there are several ways, including but not limited to:

- DESCRIBE tablename
- SHOW CREATE TABLE tablename
- SHOW COLUMNS tablename

Unfortunately, support for these varies widely between DBMSs, and is also limited by our RStudio interface (for example RSQLite, which we are using, does not support the above)

One that does work for us here through is:

```{r}
dbGetQuery(con,'
PRAGMA table_info(precipitation_full);
           ')
```

Without going into full detail (yet), the above table contains one record for each hour at each station, and contains the amount of precipitation that was measured during that hour.

Suppose that STAT 474 has been repurposed as a rain measurement corps to assist with the department that collected this data! Based on what we've covered so far, the above example, and the internet, create a table for storing the staff assignments. Table requirements:

- Everyone in the class will be holding a cup in the rain for a specific several-hour shift at a specific station; this assignment will remain the same every day
- Each person will have one off-day per week
- Each person's cup might be of a different size, measured as a float value
- The Dept. of Interior data servers can't handle the full dataset we would generate, and require a random subsample- so some people will be randomly chosen to stand in the rain without a cup. These assignments need to be recorded somehow in the table too.
- Some people in the class have [Welsh names](https://www.youtube.com/watch?v=fHxO0UdpoxM)


**Question 1:**
Type your create table statement here:

```{r}
dbGetQuery(con,'

')
```

# Part 2 - Single table queries

Run the following code to create the `Product` table in the database

```{r, message=FALSE, warning=FALSE}
sql_queries="DROP TABLE IF EXISTS product; 
CREATE TABLE product(       
       pname        VARCHAR PRIMARY KEY, -- name of the product
       price        MONEY,               -- price of the product
       category     VARCHAR,             -- category
       manufacturer VARCHAR NOT NULL     -- manufacturer
); 
INSERT INTO product values('Gizmo', 19.99, 'Gadgets', 'GizmoWorks'); 
INSERT INTO product values('PowerGizmo', 29.99, 'Gadgets', 'GizmoWorks'); 
INSERT INTO product values('MultiTouch', 203.99, 'Household', 'Hitachi'); 
INSERT INTO product values('SingleTouch', 149.99, 'Photography', 'Canon');
"
queries <- unlist(strsplit(sql_queries, ";"))
for (q in queries) dbSendQuery(con, q)
```

**Question 2:** Try writing a query to get an output table of all the products with "Touch" in the name, showing just their name and price, and sorted alphabetically by manufacturer.

Let's look at the products first:
```{r}
dbGetQuery(con,'SELECT * FROM product;')
```

Write your query here:

```{r}
dbGetQuery(con,'

')
```

Next, write a query that returns the distinct names of manufacturers that make products with "Gizmo" in the name:

```{r}
dbGetQuery(con,'

')
```

**Question 3:** Try some of these queries but first guess what they return.

```{r}
dbGetQuery(con,'
SELECT DISTINCT category FROM product ORDER BY category;
')
```

```{r}
dbGetQuery(con,'
SELECT category FROM product ORDER BY pname;
')
```

```{r}
dbGetQuery(con,'
SELECT DISTINCT category FROM product ORDER BY pname;
')
```
