---
title: "STAT 474 &ndash; Techniques for Large Data Sets"
subtitle: "Fall 2018"
date: "October 26, 2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(eval=FALSE)
library(stringr)
```

# Regular expression - Review

- Regular expression is a way to describe patterns in a string

--

- Confussing terminologies:

  - *literal:* any character used in a search or match expression
  
  - *metacharacter:* one or more special characters, have a unique meaning, and not as **literal**
  
  - *escape sequence:* a way to indicating that we want to use one of our **metacharacters** as a **literal**.
  
  - *regex string:* string that contain the regular expression.

---
# Regular expression - Review

- Some metacharacters we have seen: `.` (dot), `^` (caret), and `$` (dollar sign).

- Brackets `[ ]` match anything inside and `[^ ]` match anything that not inside

  - Can use dash for a range. Example: `[a-m0-9]` match letters from a to m and all digits.
  
  - To match the literal dash (`-`), leave it at the beginning or the end of the list.
  
- `|` is used as *alternation*. It has low precedence; usually used with the parentheses.

---
# Regular expression - Checkpoint

- Given the corpus of common words in `words`, create regular expressions that find all words that are exactly three letters long. Use argument `match = TRUE` to display only matched words.

--

- How to empirically verify the rule "i before e except after c"?

---
# Repetition

- Regex involves controlling how many times a pattern matches:

  - `?`: 0 or 1
  
  - `+`: 1 or more
  
  - `*`: 0 or more
  
```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "CC?")
str_view(x, "CC+")
str_view(x, "C[LX]+")
```

---
# Repetition

- Specify the number of matches precisely:

  - `{n}`: exactly n

  - `{n, }`: n or more

  - `{ ,m}`: at most m
  
  - `{n,m}`: between n and m

```{r}
x <- "1888 is the longest year in Roman numerals: MDCCCLXXXVIII"
str_view(x, "C{2}")
str_view(x, "C{2,}")
str_view(x, "C{2,3}")
```

--

- By default, the result will be the longest possible match. For the shortest match, put `?` after them

```{r}
str_view(x, 'C{2,3}?')
str_view(x, 'C[LX]+?')
```

---
# Repetition - Practice questions

- Describe in words what these regular expressions match: (read carefully to see if I'm using a regular expression or a string that defines a regular expression.)

  1. `^.*$`
  2. `"\\{.+\\}"`
  3. `\d{4}-\d{2}-\d{2}`
  4. `"\\\\{4}"`

--

- Create regular expressions to find all the (10-digit) phone numbers.

- Crossword game: [https://regexcrossword.com/challenges/beginner](https://regexcrossword.com/challenges/beginner)

---
# Grouping and backreferences

- Parentheses are useful to disambiguate complex expressions.

- Parentheses creates a *numbered* capturing group (number 1, 2, etc.)

  - A capturing group stores *the part of the string* matches by the part of regex inside the parentheses.

  - Can be referred to the same text as previously matched by the capturing group with **backreferences**, like `\1`, `\2`
  
```{r}
str_view(fruit, "(..)\\1", match = TRUE)
```

---
# Grouping - Practice

Describe, in words, what these expressions will match:

- `(.)\1\1`

- `"(.)(.)\\2\\1"`

- `(..)\1`

- `"(.).\\1.\\1"`

- `"(.)(.)(.).*\\3\\2\\1"`

--

*A word of caution:* because regular expressions are so powerful, it's easy to try and solve every problem with a single regular expression.

> Some people, when confronted with a problem, think “I know, I'll use regular expressions.” Now they have two problems.

---
# Detect matches

- Use function `str_detect()`

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

--

- Can take advantage of numeric conversion of logical values to do computation

```{r}
sum(str_detect(words, "^t"))
mean(str_detect(words, "[aeiou]$"))
```

- or extract the detected words with `str_subset()`

```{r}
str_subset(words, "x$")
```

---
# Extract matches

- `str_extract()` and `str_extract_all()` allow us to extract the actual text of a match.

- Example: [Harvard sentences](https://en.wikipedia.org/wiki/Harvard_sentences) used in testing of VOIP system

```{r}
length(sentences)
head(sentences)
```

--

- *Task:* Find all sentences that contain a colour.

```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

- Checkpoint: What is the usage of `color_match`?

---
# Extract matches - Example

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

--

```{r}
str_view(sentences, colour_match, match = TRUE)
```

**Question:** In the previous example, you might have noticed that the regular expression matched “reared”, which is not a colour. Modify the regex to fix the problem.

---
# Replacing matches

- `str_replace()` and `str_replace_all()` allow you to replace matches with new strings

```{r}
x <- c("apple", "pear", "banana")
str_replace(x, "[aeiou]", "-")
str_replace_all(x, "[aeiou]", "-")
```

--

- `str_replace_all()` allows multiple replacements by supplying a named vector

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

---
# Replacing matches

- Deletion is equivalent to replace a pattern with an empty string (i.e., `""`)

  - It is better to use `str_remove()` or `str_remove_all()` for deletion.

--

- `str_replacce()` allow usage of backreferences to insert components of the match

```{r}
sentences %>% 
  str_replace("([^ ]+) ([^ ]+) ([^ ]+)", "\\1 \\3 \\2") %>% 
  head(5)
```

--

**Question:** Switch the first and last letters in words. Which of those strings are still words?

---
# Splitting

- `str_split()` splits a string up into pieces.

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ")
```

- Use the argument `simplify = TRUE` to convert the results into data frame.

```{r}
sentences %>%
  head(5) %>% 
  str_split(" ", simplify = TRUE)
```

---
# Splitting words

- Instead of splitting up strings by patterns, you can also split up by character, line, sentence and word boundary:

```{r}
x <- "This is a sentence.  This is another sentence."
str_view_all(x, boundary("word"))
```

- A closer look:

```{r}
str_split(x, " ")[[1]]
str_split(x, boundary("word"))[[1]]
```

---
# Miscellaneous 

- To match a specific sequence of character, use `fixed()` to speed up the performance (up to 3x faster than the regular expression)

```{r}
start_time <- Sys.time()
str_view(sentences, fixed("the"), match = TRUE)
Sys.time() - start_time
```

- `fixed()` doesn't accept the special regular expressions. 

--

- `str_locate()` and `str_locate_all()` give you the starting and ending positions of each match.

- `str_to_lower()` and `str_to_upper()` convert letters in strings into their lower- and uppercases, respectively.

- `str_trim()` removes space in front and at the end of the string

- `str_squish()`reduces repeated whitespace inside a string.